---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# R Youth Homelessness Analyzer *(RYHA)*

<!-- badges: start -->
<!-- badges: end -->

The goal of `{ryha}` is to provide a web-based application for in-depth, user-friendly consumption of Youth Homelessness data across the State of Ohio.

## Installation

This Shiny application is built using the [{golem} framework](https://thinkr-open.github.io/golem/), which means that this app is actually an *R package*.

You can install the development version of `{ryha}` from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("ketchbrookanalytics/ryha")
```

## Running the App

After installing the package, you can launch the app using the package's `run_app()` function.

```{r run_app, eval = FALSE}
ryha::run_app()
```

## Data

### Definitions

* *"Youth"* is defined as any person within the program who is age 24 or younger.

### Source

The data within this application are collected by each grantee in a format compliant with the *HMIS Data Standards*.

The data dictionary for each element across each file can be found here: 
[https://files.hudexchange.info/resources/documents/FY-2022-HMIS-Data-Dictionary.pdf](https://files.hudexchange.info/resources/documents/FY-2022-HMIS-Data-Dictionary.pdf)

### Methodology for Analysis

This app gives users the ability to slice & filter the data across many different demographic variables and attributes about each individual in the program. These attributes are grouped into the following categories within the app:

```{r data_sources_tbl, echo = FALSE}

tibble::tribble(
  ~Attribute, ~`Source File`, 
  "Age", "*Client.csv*", 
  "Ethnicity", "*Client.csv*", 
  "Gender", "*Client.csv*", 
  "Military History", "*Client.csv*", 
  "Disabilities", "*Disabilities.csv*", 
  "Education", "*EmploymentEducation.csv*", 
  "Employment", "*EmploymentEducation.csv*", 
  "Living Situation", "*Enrollment.csv*", 
  "Welfare History", "*Enrollment.csv*", 
  "Program Exit Status", "*Exit.csv*", 
  "Trafficking", "*Exit.csv*", 
  "Domestic Violence History", "*HealthAndDV.csv*", 
  "Overall Health (General, Dental, and Mental)", "*HealthAndDV.csv*", 
  "Pregnancy Status", "*HealthAndDV.csv*", 
  "Income", "*IncomeBenefits.csv*", 
  "Unemployment & Other Benefits Received", "*IncomeBenefits.csv*", 
  "Insurance", "*IncomeBenefits.csv*"
  
) |> 
  knitr::kable(caption = "*Table 1*")

```

### Historical Database

Privileged users have access to historical data within the application, which gives the ability to view trends over time across metrics.

### Individual Program File Upload

Non-privileged users have the ability to upload their own (individual program) data into the app for analysis.

#### File Storage

Users can upload one Zip file containing *.csv* files representing a single export of HMIS data for their individual program (i.e., grantee). The uploaded data will pass through a data processing script, which will break the *.csv* files up into smaller, more modular datasets that align with the attributes described in [Table 1](#methodology-for-analysis), and stored as *.parquet* files in the data lake.

The data lake is partitioned by **program (grantee)** and **export date range**. An example of the data lake directory structure across two example programs (and three example files) can be seen below:

```{r data_lake_structure, echo = FALSE}
fs::dir_tree("example_data_lake")
```

